### 2013-06-07 KapowskiChronicles/conclusions.tex
ENH:  Adding conclusion section.

\section{Discussion and Conclusions}

Imaging biomarkers such as cortical thickness play an 
important role in neuroscience research.  In this work we 
detailed our open source strategy for estimating
cortical thickness directly from T1 images and demonstrated
its utility on a large collection of public brain data from
multiple databases acquired at multiple sites.  To our knowledge
this cohort constitutes the largest collection of cortical
thickness data processed in a single study.

In this work, we only explored a portion of the potentially
interesting investigations possible with these data.  However,
since all these data are publicly available, further work can
be easily pursued by us or even other interested groups.  We 
note that in addition to making the scripts, csv result files,
and cortical thickness maps available to researchers, 
posterior probability images are also offered permitting additional
quantification such as voxel-based morphometry \cite{ashburner2000}.